Global code motion of congruent instructions

### 2017-11-08 cgo/paper.tex
move barriers above

\subsubsection{Legality of movable instructions}
\label{subsec:legality}
Since the equality of candidates is purely based on the value numbers, we also
need to establish if hoisting them or sinking them would be legal. The
completeness of CHI is a necessary condition for hoistability but not
sufficient. This is because congruence of value numbers only implies that they
compute same value if the inputs are same. So for instructions like memory
references, they may compute different values should the memory get modified
along the path. For that we need to check for intersecting side-effects along
the path.

For hoisting, we need to prove that on all execution paths from the hoisting
point to the end of the program the side effects appear in the same order. In
general, we cannot introduce a new computation along any path of execution
without checking for undefined behavior \cite{undef}. For that we need to see a)
if I aliases with any instruction before it in its basic block and b) Walk the
CFG from its basic block 'upwards' to its PDF and look for any interfering
side-effects. If there aren't any then it is safe to hoist.

Subsequently, it is checked that the side-effects of the computations (if any)
do not intersect (alias) with any side-effects between the instructions to be
moved and their hoisting/sinking point. It is also necessary to check if there
are indirect branch targets, e.g., landing pad, case statements, goto labels,
etc., along the path because it becomes difficult to prove safety checks in
those cases. In our current implementation we discard candidates on those paths.

@@

A general global scheduling algorithm also requires checks for undefined values
when introducing a new computation along a path. Since only very busy
expressions are moved in the current implementation, there is no need to check
for undefinedness resulting due to movement of instructions. This simplifies the
implementation.

@@

\subsubsection{Barriers}
\label{subsec:barriers}
Barriers are based on the concept of pinned instructions \cite{click1995global}
but extended to adapt to \LLVM{} IR. A basic-block in \LLVM{} IR is actually an
extended basic-block because there might be non-returning calls in the middle of
a basic-block. Essentially, any instruction that cannot guarantee progress is
marked as a barrier. In the absence of context, as in our current
implementation, some instructions which might be safe are still classified as
barriers, e.g., volatile loads/stores, calls with missing attributes.

>>>>

Before we consider factors affecting the legality of \GCM{}, we introduce the
concept of barriers.

\subsection{Barriers}
\label{subsec:barriers}
Barriers are based on the concept of pinned instructions \cite{click1995global}
but extended to adapt to \LLVM{} IR. A basic-block in \LLVM{} IR is actually an
extended basic-block because there might be non-returning calls in the middle of
a basic-block. Essentially, any instruction that cannot guarantee progress is
marked as a barrier. In the absence of context, as in our current
implementation, some instructions which might be safe are still classified as
barriers, e.g., volatile loads/stores, calls with missing attributes.

Computing barriers allows for efficient removal of non-movable candidates, this
makes the legality check converge faster.

\subsubsection{Legality of movable instructions}
\label{subsec:legality}
Since the equality of movable candidates is purely based on the value numbers,
we also need to establish if hoisting them or sinking them would be legal. The
completeness of CHI is a necessary condition for hoistability but not
sufficient. This is because congruence of value numbers only implies that they
compute same value if the inputs are same. So for instructions like memory
references, they may compute different values should the memory get modified
along the path. For that we need to check for intersecting side-effects along
the path.

For hoisting, we need to prove that on all execution paths from the hoisting
point to the end of the program the side effects appear in the same order. In
general, we cannot introduce a new computation along any path of execution
without checking for undefined behavior \cite{undef}.  For that, it is checked
that the side-effects of the computations (if any) do not intersect (alias) with
any side-effects between the instructions to be moved and their hoisting/sinking
point. It is also necessary to check if there are indirect branch targets, e.g.,
landing pad, case statements, goto labels, etc., along the path because it
becomes difficult to prove safety checks in those cases. In our current
implementation we discard candidates on those paths.

@@

@@
